### 2025-04-27 paper.tex
Layout final paper

\documentclass[conference]{IEEEtran}
\IEEEoverridecommandlockouts
% The preceding line is only needed to identify funding in the first footnote. If that is unneeded, please comment it out.
\usepackage{cite}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{algorithmic}
\usepackage{url}
\usepackage{graphicx}
\usepackage{textcomp}
\usepackage{xcolor}
\usepackage{tikz}
\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}
\begin{document}
\title{Formally Verifying Voter Security Properties in DAO Smart Contracts}

\author{
    \makebox[\textwidth][c]{%
        \begin{minipage}{0.45\textwidth}
            \centering
            \textbf{Anish Toomu} \\
            \textit{Department of Computer Science} \\
            \textit{University of North Carolina at Chapel Hill} \\
            Chapel Hill, North Carolina, USA
        \end{minipage}
        \hfill
        \begin{minipage}{0.45\textwidth}
            \centering
            \textbf{Aidan Maguire} \\
            \textit{Department of Computer Science} \\
            \textit{University of North Carolina at Chapel Hill} \\
            Chapel Hill, North Carolina, USA
        \end{minipage}
    }\\[1.5em]

    \makebox[\textwidth][c]{%
        \begin{minipage}{0.6\textwidth}
            \centering
            \textbf{Jeffrey Zhang} \\
            \textit{Department of Computer Science} \\
            \textit{University of North Carolina at Chapel Hill} \\
            Chapel Hill, North Carolina, USA
        \end{minipage}
    }
}

\maketitle

\section{Abstract}
To be written

\section{Introduction}
This project aims to make use of two existing security tools to define and verify security properties for voting methods across a suite of Ethereum blockchain programs. Many Ethereum blockchain programs take the form of a DAO, or Decentralized Autonomous Organization, wherein users holding tokens that verify them as members of the organization can spend those tokens to vote on decision-making for the DAO as a whole. There are several types of vulnerabilities that have been identified in these types of contracts, but for this paper we will be looking at vulnerabilities surrounding the vote system that allow attackers to exploit the democratic principles behind DAOs.

Our proposal intends to formally verify a suite of the most used DAO contracts to confirm that they are secure. We will require the use of Scribble, a specification language that allows us to define security properties for DAO contracts as annotations within the contracts themselves. We will then use Mythril, a security-analysis tool, to verify each of our newly defined security properties for our suite of popular contracts. If successful, we will increase the confidence in these widely-used contracts. Possible challenges include the properties not being entirely specifiable in Scribble's specification language, and some contracts being too complex to analyze with Mythril due to state explosion.

\section{State of the Art}

DAOs are a type of smart contract that define the rules and structure of a organization. In more general terms, DAOs represent decentralized communities that can allocate resources without centralized leadership. While there exist tools for various methods of verifying and bug-finding in smart contracts including static analysis[1], model checking[2], and symbolic execution[3], they haven’t been narrowed down to focus on DAOs, and the specific exploits that come with their voting systems, a key, shared property across DAOs and a critical part of their operations. We have chosen to use an existing symbolic execution-based tool, Mythril[4], along with Scribble[5], a specification language for properties in this work.

\section{Background}
To understand our proposed enhancements to Mythril, we must first grasp the fundamental concepts of the blockchain, the Ethereum blockchain and its innovative smart contracts, and finally DAOs, a subset of smart contracts with common functionalities.

\subsection{Blockchain}
A blockchain serves as a ledger of transactions that is transparent, decentralized, and secure. Each block in the chain contains information about transactions allowing anyone to inspect them at any time. Additionally, each block contains a cryptographic hash of the previous block in the chain. This chain means that in the event of a block being tampered with by an attacker, all of the altered blocks' successors would also be changed. The consensus needed to accept a chain also provides further security. If the majority of the computational power on the network is controlled by honest parties then the honest chain will be accepted. An attacker seeking to change a block would need to alter the block, redo all the proof-of-work computation for it and its successor blocks, and then catch up to the honest chain to be accepted. This proves to be computationally infeasible[6].

\subsection{Ethereum and Smart Contracts}
Ethereum, as proposed by Vitalik Buterin in 2014, seeks to provide an alternative to Bitcoin with a focus on the development of decentralized applications. This is achieved through a blockchain that contains a built-in Turing-complete language[7]. This language allows users to write smart contracts that are stored on the blockchain. Smart contracts are a type of account, they have a balance and can be interacted with through transactions. They differ from users’ externally owned accounts in their method of control. Where externally owned accounts are controlled by user decisions, smart contracts are programs that run on the blockchain according to their code. Users or other smart contracts interact with smart contracts by sending a transaction that executes the functions defined by the contract’s code. Smart contracts automatically execute when their conditions are met, are universally visible, and are unable to be changed once uploaded[8]. In this way smart contracts provide a trustworthy, automated way to enforce agreements, reducing the need for third-party enforcement, minimizing exceptions, and lowering the risk of fraud[9].

\subsection{Decentralized Autonomous Organizations}
Decentralized Autonomous Organizations (DAOs) use the blockchain and smart contracts as a way to govern and operate an organization without a central authority. DAOs allow participants to vote on proposed policy changes, actions taken by the organization, and any number of other decisions. These rules are automatically enforced by smart contracts. The first notable DAO, “The DAO” was formed on the Ethereum blockchain in 2016 as a venture capital fund. Users could exchange ether for tokens that gave holders voting power in The DAO proportional to the amount of tokens owned[10]. The organization managed to crowdfund \$168M worth of ether representing one of the largest projects on the Ethereum blockchain. The project is now synonymous with a programming vulnerability in the organization's smart contract that allowed an attacker to siphon off \$50M from The DAO’s[11] gathered funds. This resulted in the controversial hard fork of the Ethereum blockchain and the shuttering of the organization. This example highlights the critical nature of DAO smart contract security as they manage large amounts of currency and interact with a large number of users. Today, a great variety of DAOs exist on Ethereum ranging from communities of creators to venture capital groups[12].

\section{Threat Model}
DAO contract voting vulnerabilities refers to a class of vulnerabilities where an attacker exploits the contract voting mechanism to vote twice with the same tokens, remove other people's votes, vote on a proposal and immediately put it into effect without giving other token holders a chance to vote themselves, or other similar exploits that undermine the decentralized democratic goals of DAO contracts. In our analysis, we assume that attackers have complete visibility of the contract code, and the ability to deploy their own contracts to the blockchain that interface with DAO contracts[13]; however, we are assuming that attackers cannot launch larger Ethereum network attacks such as network congestion or miner manipulation that would require greater coordination and resources. We consider nonvoting-related attacks such as integer overflow, reentrancy, and gas DoS attacks as out of scope for this paper.

\section{Design}
\subsection{Defining security properties for DAO smart contracts}
\begin{center}
\begin{tikzpicture}[scale=0.15]
\tikzstyle{every node}+=[inner sep=0pt]
\draw [black] (9,-32.6) circle (3);
\draw (9,-32.6) node {$s_1,s_2$};
\draw [black] (23.6,-16) circle (3);
\draw (23.6,-16) node {$D_1$};
\draw [black] (23.6,-49.3) circle (3);
\draw (23.6,-49.3) node {$W_1$};
\draw [black] (42.4,-16) circle (3);
\draw (42.4,-16) node {$D_2$};
\draw [black] (57.2,-16) circle (3);
\draw (57.2,-16) node {$D_3$};
\draw [black] (43.6,-49.3) circle (3);
\draw (43.6,-49.3) node {$W_2$};
\draw [black] (57.2,-49.3) circle (3);
\draw (57.2,-49.3) node {$W_3$};
\draw [black] (46.4,-32.6) circle (3);
\draw (46.4,-32.6) node {$I$};
\draw [black] (28.7,-32.6) circle (3);
\draw (28.7,-32.6) node {$R$};
\draw [black] (10.98,-30.35) -- (21.62,-18.25);
\fill [black] (21.62,-18.25) -- (20.71,-18.52) -- (21.47,-19.18);
\draw (16.84,-25.75) node [right] {$a_1$};
\draw [black] (10.97,-34.86) -- (21.63,-47.04);
\fill [black] (21.63,-47.04) -- (21.48,-46.11) -- (20.72,-46.77);
\draw (16.84,-39.5) node [right] {$a_2$};
\draw [black] (26.6,-16) -- (39.4,-16);
\fill [black] (39.4,-16) -- (38.6,-15.5) -- (38.6,-16.5);
\draw (33,-16.5) node [below] {$c_1$};
\draw [black] (45.4,-16) -- (54.2,-16);
\fill [black] (54.2,-16) -- (53.4,-15.5) -- (53.4,-16.5);
\draw (49.8,-16.5) node [below] {$d_1$};
\draw [black] (26.6,-49.3) -- (40.6,-49.3);
\fill [black] (40.6,-49.3) -- (39.8,-48.8) -- (39.8,-49.8);
\draw (33.6,-49.8) node [below] {$c_2$};
\draw [black] (46.6,-49.3) -- (54.2,-49.3);
\fill [black] (54.2,-49.3) -- (53.4,-48.8) -- (53.4,-49.8);
\draw (50.4,-49.8) node [below] {$d_2$};
\draw [black] (8.934,-29.602) arc (-182.00565:-319.9871:26.325);
\fill [black] (8.93,-29.6) -- (9.46,-28.82) -- (8.46,-28.79);
\draw [black] (55.065,-51.406) arc (-48.49526:-169.72438:27.748);
\fill [black] (9.37,-35.58) -- (9.03,-36.45) -- (10.01,-36.27);
\draw [black] (43.4,-32.6) -- (31.7,-32.6);
\fill [black] (31.7,-32.6) -- (32.5,-33.1) -- (32.5,-32.1);
\draw (37.55,-33.1) node [below] {$f$};
\draw [black] (24.48,-18.87) -- (27.82,-29.73);
\fill [black] (27.82,-29.73) -- (28.06,-28.82) -- (27.11,-29.11);
\draw (25.38,-24.93) node [left] {$b_1$};
\draw [black] (24.48,-46.43) -- (27.82,-35.47);
\fill [black] (27.82,-35.47) -- (27.11,-36.09) -- (28.07,-36.38);
\draw (26.92,-41.57) node [right] {$b_2$};
\draw [black] (43.1,-18.92) -- (45.7,-29.68);
\fill [black] (45.7,-29.68) -- (46,-28.79) -- (45.02,-29.02);
\draw (43.64,-24.73) node [left] {$e_1$};
\draw [black] (44.1,-46.34) -- (45.9,-35.56);
\fill [black] (45.9,-35.56) -- (45.28,-36.27) -- (46.26,-36.43);
\draw (45.71,-41.18) node [right] {$e_2$};
\end{tikzpicture}
\end{center}

\section{Implementation}
To be written

\section{Evaluation}

This project investigates whether symbolic execution can be used to verify the safety and liveness properties of DAO voting functions. It also explores the scalability of this approach when applied to more complex voting mechanisms and assesses whether the method can be adapted to different DAOs' smart contracts. If automation isn't feasible, we want to investigate the manual effort required for adaptation. Our experiments will involve both toy examples and open-source smart contracts, focusing on analyzing DAO voting logic and defining the properties needed to verify it.

\section{Conclusion}
To be written

\section{Who Did What}
\begin{itemize}
    \item Aidan: Contributed to the Introduction, Threat Model, and References
    \item Anish: Contributed to the Approach, Set up repository, dependencies, and starter DAO contract/annotations
    \item Jeffrey: Contributed to the Background, State of the Art, and Evaluation sections
\end{itemize}

\begin{thebibliography}{00}
\bibitem{b1} J. Feist, G. Grieco, and A. Groce, “Slither: A static analysis framework for smart contracts,” 2019, pp. 8–15. doi: https://doi.org/10.1109/WETSEB.2019.00008. Available: \url{https://ieeexplore.ieee.org/abstract/document/8823898?casa_token=SNthQs5W2d0AAAAA:XwX332JMCkj3hw8XZd9nKBxe2bKJZ3v3yUNJ3jPtangZ3L2nISxh_rbY8fBW2zXkaWa0QnqgqVE}
\bibitem{b2} Z. Nehaï, P.-Y. Piriou, and F. Daumas, “Model-checking of smart contracts,” 2018, pp. 980–987. Available: \url{https://ieeexplore.ieee.org/abstract/document/8726806}
\bibitem{b3} J. He, M. Balunović, N. Ambroladze, P. Tsankov, and M. Vechev, “Learning to Fuzz from Symbolic Execution with Application to Smart Contracts,” Proceedings of the 2019 ACM SIGSAC Conference on Computer and Communications Security, pp. 531–548, Nov. 2019, Available: \url{https://dl.acm.org/doi/abs/10.1145/3319535.3363230?casa_token=5cBcZneJt7oAAAAA:sPluGu-gtO_db3qLqHUcylo4ksy4Rx0FBkiEBFom7You4VltkgLyZanIwgg-WHfLDVmrwmU0MbdpGA}
\bibitem{b4} N. Sharma and S. Sharma, “A Survey of Mythril, A Smart Contract Security Analysis Tool for EVM Bytecode,” Indian Journal of Natural Sciences, vol. 12, no. 75, Dec. 2022, Available:
\url{https://www.researchgate.net/profile/Swati-Sharma-171/publication/366391033_A_Survey_of_Mythril_A_Smart_Contract_Security_Analysis_Tool_for_EVM_Bytecode/links/639ecbdc095a6a77743c8073/A-Survey-of-Mythril-A-Smart-Contract-Security-Analysis-Tool-for-EVM-Bytecode.pdf}
\bibitem{b5} ConsenSysDiligence, “GitHub - ConsenSysDiligence/scribble: Scribble instrumentation tool,” GitHub, Apr. 09, 2025. Available:  
\url{ConsenSysDiligence, “GitHub - ConsenSysDiligence/scribble: Scribble instrumentation tool,” GitHub, Apr. 09, 2025. https://github.com/ConsenSysDiligence/scribble/tree/develop (accessed Apr. 10, 2025).}
\bibitem{6} S. Nakamoto, “Bitcoin: a Peer-to-Peer Electronic Cash System,” bitcoin.org, Oct. 2008. Available: 
\url{https://bitcoin.org/bitcoin.pdf}
\bibitem{7} V. Buterin, “Ethereum Whitepaper,” Ethereum, 2014. Available: 
\url{https://ethereum.org/en/whitepaper/}
\bibitem{8} P. Wackerow, “Introduction to smart contracts,” ethereum.org, Sep. 02, 2022. Available: 
\url{https://ethereum.org/en/developers/docs/smart-contracts/}
\bibitem{9} N. Szabo, “Smart Contracts,” www.fon.hum.uva.nl, 1994. 
\url{https://www.fon.hum.uva.nl/rob/Courses/InformationInSpeech/CDROM/Literature/LOTwinterschool2006/szabo.best.vwh.net/smart.contracts.html}
\bibitem{10} C. Jentzsch, “DECENTRALIZED AUTONOMOUS ORGANIZATION TO AUTOMATE GOVERNANCE,” 2016. Available: 
\url{https://lawofthelevel.lexblogplatformthree.com/wp-content/uploads/sites/187/2017/07/WhitePaper-1.pdf}
\bibitem{11} C. Shier et al., “Understanding a Revolutionary and Flawed Grand Experiment in Blockchain: The DAO Attack,” SSRN Electronic Journal, 2017, doi: 
\url{https://doi.org/10.2139/ssrn.3014782.}
\bibitem{12} “List of 41 DAOs (2025),” Alchemy, 2025. Available:
\url{https://www.alchemy.com/dapps/top/daos}
\bibitem{13} K. Nekrasov, “DAO Voting Vulnerabilities,” Mixbytes.io, 2024.
\url{https://mixbytes.io/blog/dao-voting-vulnerabilities}
\bibitem{14} HausDAO, “GitHub - HausDAO/Molochv2.1: Moloch DAO v2 with multi-summoner capabilities, plus register function for metadata.,” GitHub, 2020. Available:
\url{https://github.com/HausDAO/Molochv2.1}
\end{thebibliography}
\end{document}